### 2020-07-28 manuscript/manuscript.Rmd
hello world

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

csl               : "apa.csl"
documentclass     : "apa7"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
# Load packages
library(papaja)
library(reticulate)
library(MASS)
library(Matrix)
library(lme4)
library(buildmer)
library(emmeans)
library(magrittr)
library(tidyverse)

# Chunk options
knitr::opts_chunk$set(include = FALSE, echo = FALSE)

# Specify conda environment for Python
use_condaenv("r-reticulate", required = TRUE)

# Bibliography file
r_refs("manuscript/r-references.bib")
```

```{python exp1-preprocessing}
# Import libraries
import mne
import glob
import os

# Define parameters for preprocessing
preproc_params = dict(n_components=15, method='fastica',           # ICA parameters
                      l_freq=0.1, h_freq=30,                       # Filter edges
                      event_id={'match': 221, 'mismatch': 222},    # EEG triggers
                      tmin=-0.5, tmax=1.498,                       # Length of epochs
                      baseline=(-0.2,0))                           # Baseline correction
reject = dict(eeg=200e-6)                                          # Rejection threshold

# # Debug
# locals().update(preproc_params)
# vhdr_fname = 'data/exp1/EEG/Vp0001.vhdr'

# Define preprocessing function
def preproc(vhdr_fname, n_components, method, l_freq, h_freq, event_id, tmin, tmax, baseline):
  # Load EEG data
  raw = mne.io.read_raw_brainvision(vhdr_fname, preload=True)
  # Create new virtual EOG channels
  raw = mne.set_bipolar_reference(raw, 'Auge_u', 'Fp1', ch_name='VEOG', drop_refs=False)
  raw = mne.set_bipolar_reference(raw, 'F9', 'F10', ch_name='HEOG', drop_refs=False)
  raw.set_channel_types(mapping={'VEOG': 'eog', 'HEOG': 'eog'})
  # Add EasyCap electrode layout, removing any excessive channels
  montage = mne.channels.make_standard_montage('easycap-M1')
  raw.drop_channels(list(set(raw.ch_names) - set(montage.ch_names) - set(['VEOG', 'HEOG'])))
  raw.set_montage(montage=montage)
  # Re-reference to common average
  raw, _ = mne.set_eeg_reference(raw, 'average')
  # Run ICA on a copy of the data
  filt_raw = raw.copy()
  filt_raw.load_data().filter(l_freq=1, h_freq=None)
  ica = mne.preprocessing.ICA(n_components=n_components, random_state=12345, method=method)
  ica.fit(filt_raw)
  # Remove bad components based on correlations with EOG
  eog_indices, eog_scores = ica.find_bads_eog(raw)
  ica.exclude = eog_indices
  raw = ica.apply(raw)
  # Apply band-pass filter
  raw = raw.filter(l_freq=l_freq, h_freq=h_freq)
  # Epoching including baseline correction
  events, _ = mne.events_from_annotations(raw, verbose=False)
  epochs = mne.Epochs(raw, events=events, event_id=event_id, tmin=tmin, tmax=tmax,
                      baseline=baseline, preload=True)

  # # Autoreject                    
  # import autoreject
  # 
  # ar = autoreject.AutoReject(thresh_method='random_search', verbose=False)
  # epochs_clean = ar.fit_transform(epochs)
  # 
  # rejected = epochs_clean.drop_log
  # rejected[:] = [item for item in rejected if item != ['IGNORED']]
  # rejected = [i for i in range(len(rejected)) if rejected[i] != []]
  # 
  # ransac = autoreject.Ransac(verbose=False)
  # epochs = ransac.fit_transform(epochs)
  # 
  # reject = autoreject.get_rejection_threshold(epochs)

  return epochs

# List of raw EEG filenames
fnames_exp1 = sorted(glob.glob('data/exp1/EEG/*.vhdr'))

# Check if preprocessing was done already (delete the file to re-run)
if os.path.exists('output/exp1-epo.fif'):
  # Load preprocessed data from file
  epochs_exp1 = mne.read_epochs('output/exp1-epo.fif', preload=True)
else:
  # Actually apply the preprocessing function
  epochs_exp1 = [preproc(fname, **preproc_params) for fname in fnames_exp1]
  # Combine into a single data set
  epochs_exp1 = mne.concatenate_epochs(epochs_exp1)
  # Backup epochs to the export folder
  epochs_exp1.save('output/exp1-epo.fif')

# Get indices of bad epochs despite ICA
rej_exp1 = epochs_exp1.copy().drop_bad(reject=reject).drop_log
rej_exp1[:] = [item for item in rej_exp1 if item != ['IGNORED']]
rej_exp1 = [i for i in range(len(rej_exp1)) if rej_exp1[i] != []]
```

```{r exp1-analysis}
# Define function for processing behavioral data
preproc_behav <- function(txt_fname){
  # Read log file
  dat <-  suppressMessages(suppressWarnings(read_tsv(txt_fname)))
  # Remove filler stimuli (well-known objects)
  dat %<>% filter(bek_unbek != "bekannt")
  # Add a column for the part of the experiment
  dat %<>% mutate(part = as_factor(Wdh - 210))

  # Assign stimuli to conditions based on keywords and button presses
  known <- dat %>% filter(part == "1" & Tastencode == 201) %>% pull(StimID)
  insight <- dat %>%filter(part == "2" & Bed == "richtig" & Tastencode %in% c(201, 202)) %>% pull(StimID)
  naive <- dat %>% filter(part == "2" & Bed == "falsch" & Tastencode %in% c(203, 204)) %>% pull(StimID)
  dat %<>% mutate(condition = as_factor(case_when(StimID %in% known ~ "Known",
                                                  StimID %in% insight ~ "Insight",
                                                  StimID %in% naive ~ "Naive")))
  # Remove unrealistically short RTs
  dat %<>% mutate(RT = replace(RT, RT < 200, NA))
  return(dat)}

# List of filenames for behavioral data
fnames_exp1 <- list.files("data/exp1/RT", pattern = ".txt", full.names = TRUE)

# Actually apply the preprocessing function
data_exp1 <- map_dfr(fnames_exp1, preproc_behav)

# Add factorized columns for participants and items (the random variables)
data_exp1 %<>% mutate(participant = as_factor(VPNummer), item = as_factor(StimID))

# Tabulate percentage of stimuli per condition
stims_exp1 <- table(data_exp1$participant, data_exp1$condition)
stims_exp1 <- colMeans(stims_exp1)
stims_exp1["Known"]/(3*120) # Three repititions and 120 objects
stims_exp1[c("Insight", "Naive")]/(3*60) # Three repitions and 60 objects each

# Import epochs from Python, converting Volts to Microvolts
epochs_exp1 <- py$epochs_exp1$get_data()*1e6

# Set rejected epochs to NA
rej_exp1 <- py$rej_exp1 + 1
epochs_exp1[rej_exp1,,] <- NA

# Create vectors with sample time points (in ms) and electrodes
tmin <- py$preproc_params$tmin*1000
tmax <- py$preproc_params$tmax*1000
sfreq <- py$epochs_exp1$info['sfreq']
times <- seq(tmin, tmax, 1000/sfreq)
els <- py$epochs_exp1$ch_names

# Define ERP components of interest
comps <- tibble(name = c("P1", "N1", "N400"),
                start = c(100, 150, 400),
                stop = c(150, 200, 700),
                roi = list(c("PO3", "PO4", "POz", "O1", "O2", "Oz"),
                           c("P7", "P8", "PO7", "PO8", "PO9", "PO10"),
                           c("C1", "C2", "Cz", "CP1", "CP2", "CPz")))

# Define function to compute single-trial ERPs
compute_erps <- function(epochs, els, name, start, stop, roi){
  erps <- epochs[, which(els %in% roi), which(times %in% start:stop)]
  erps <- apply(erps, 1, mean, na.rm = TRUE)
  return(erps)}

# Apply the function and bind ERPs to the behavioral data
data_exp1 <- pmap_dfc(comps, compute_erps, epochs = epochs_exp1, els = els) %>% set_names(comps$name) %>% cbind(data_exp1, .)

# Remove trials with missing ERPs or RTs
rej_exp1_erp <- data_exp1 %>% filter(is.na(P1)) %>% count(participant, .drop = FALSE) %>% pull(n)
data_exp1 %<>% filter(!is.na(P1))
rej_exp1_rt <- data_exp1 %>% filter(is.na(RT)) %>% count(participant, .drop = FALSE) %>% pull(n)
data_exp1 %<>% filter(!is.na(RT))

# Remove any trials that don't beling to either of the two conditions
data_exp1 %<>% filter(condition != "Known") %>% droplevels()

# Set sliding difference coding for condition (insight-naive) and part (2-1, 3-1)
t(contrasts.condition <- t(cbind(c("Insight" = 1, "Naive" = -1))))
t(contrasts.part <- t(cbind(c("1" = -1, "2" = 1, "3" = 0),
                            c("1" = 0, "2" = -1, "3" = 1))))
contrasts(data_exp1$condition) <- ginv(contrasts.condition)
contrasts(data_exp1$part) <- ginv(contrasts.part)

# Parameters for linear mixed-effects models (LMMs)
form <- tabulate.formula(~ part*condition + (part*condition|participant) + (part*condition|item))
form %<>% mutate(block = replace(block, is.na(grouping), "fixed"))
ctrl_params <- lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))

# Parameters for follow-up contrasts
specs = pairwise ~ condition|part

# Define function to compute LMMs and contrasts
compute_models <- function(dep, formula, data, control, specs){
  build <- buildmer(dep = dep, formula = formula, data = data, ddf = "Satterthwaite", calc.anova = TRUE, control = control)
  emm_options(lmer.df = "Satterthwaite", lmerTest.limit = Inf)
  conts <- emmeans(build@model, specs = specs, infer = TRUE)
  mod <- list("model" = build@model, "anova" = build@anova, "summary" = build@summary, "contrasts" = conts)
  return(mod)}

# Check if models were computed already (delete the file to re-run)
if (file.exists("output/exp1-stats.RDS")){
   # Load models from file
  models_exp1 <- readRDS("output/exp1-stats.RDS")
} else {
  # Actually apply the modelling function
  models_exp1 <- map(comps$name, compute_models, formula = form, data = data_exp1, control = ctrl_params, specs = specs) %>%
    set_names(comps$name)
  # Backup epochs to the export folder
  saveRDS(models_exp1, "output/exp1-stats.RDS")}
```

```{python exp2-preprocessing}
# List of raw EEG filenames
fnames_exp2 = sorted(glob.glob('data/exp2/EEG/*.vhdr'))

# Check if preprocessing was done already (delete the file to re-run)
if os.path.exists('output/exp2-epo.fif'):
  # Load preprocessed data from file
  epochs_exp2 = mne.read_epochs('output/exp2-epo.fif', preload=True)
else:
  # Actually apply the preprocessing function
  epochs_exp2 = [preproc(fname, **preproc_params) for fname in fnames_exp2]
  # Combine into a single data set
  epochs_exp2 = mne.concatenate_epochs(epochs_exp2)
  # Backup epochs to the export folder
  epochs_exp2.save('output/exp2-epo.fif')

# Get indices of bad epochs despite ICA
rej_exp2 = epochs_exp2.copy().drop_bad(reject=reject).drop_log
rej_exp2[:] = [item for item in rej_exp2 if item != ['IGNORED']]
rej_exp2 = [i for i in range(len(rej_exp2)) if rej_exp2[i] != []]
```

```{r exp2-analysis}
# List of filenames for behavioral data
fnames_exp2 <- list.files("data/exp2/RT", pattern = ".txt", full.names = TRUE)

# Actually apply the preprocessing function
data_exp2 <- map_dfr(fnames_exp2, preproc_behav)

# Add factorized columns for participants and items (the random variables)
data_exp2 %<>% mutate(participant = as_factor(VPNummer), item = as_factor(StimID))

# Tabulate percentage of stimuli per condition
stims_exp2 <- table(data_exp2$participant, data_exp2$condition)
stims_exp2 <- colMeans(stims_exp2)
stims_exp2["Known"]/(3*120) # Three repititions and 120 objects
stims_exp2[c("Insight", "Naive")]/(3*60) # Three repitions and 60 objects each

# Import epochs from Python, converting Volts to Microvolts
epochs_exp2 <- py$epochs_exp2$get_data()*1e6

# Set rejected epochs to NA
rej_exp2 <- py$rej_exp2 + 1
epochs_exp2[rej_exp2,,] <- NA

# Apply function to compute single-trial ERPs and to the behavioral data
data_exp2 <- pmap_dfc(comps, compute_erps, epochs = epochs_exp2, els = els) %>% set_names(comps$name) %>% cbind(data_exp2, .)

# Remove trials with missing ERPs or RTs
rej_exp2_erp <- data_exp2 %>% filter(is.na(P1)) %>% count(participant, .drop = FALSE) %>% pull(n)
data_exp2 %<>% filter(!is.na(P1))
rej_exp2_rt <- data_exp2 %>% filter(is.na(RT)) %>% count(participant, .drop = FALSE) %>% pull(n)
data_exp2 %<>% filter(!is.na(RT))

# Remove any trials that don't beling to either of the two conditions
data_exp2 %<>% filter(condition != "Known") %>% droplevels()

# Set sliding difference coding for condition (insight-naive) and part (2-1, 3-1)
contrasts(data_exp2$condition) <- ginv(contrasts.condition)
contrasts(data_exp2$part) <- ginv(contrasts.part)

# Check if models were computed already (delete the file to re-run)
if (file.exists("output/exp2-stats.RDS")){
   # Load models from file
  models_exp2 <- readRDS("output/exp2-stats.RDS")
} else {
  # Actually apply the modelling function
  models_exp2 <- map(comps$name, compute_models, formula = form, data = data_exp2, control = ctrl_params, specs = specs) %>%
    set_names(comps$name)
  # Backup epochs to the export folder
  saveRDS(models_exp2, "output/exp2-stats.RDS")}
```

```{r joint-analysis}
# Combine data from Experiments 1 and 2 (requires changing the participant IDs for Experiment 2)
data_joint <- data_exp2 %>%
  mutate(participant = fct_relabel(participant, ~ paste0(., "_2"))) %>%
  bind_rows(data_exp1, ., .id = "experiment") %>%
  mutate(experiment = as_factor(experiment))

# Set sliding difference coding for experiment (2-1), condition (insight-naive), and part (2-1, 3-1)
t(contrasts.experiment <- t(cbind(c("1" = -1, "2" = 1))))
contrasts(data_joint$experiment) <- ginv(contrasts.experiment)
contrasts(data_joint$condition) <- ginv(contrasts.condition)
contrasts(data_joint$part) <- ginv(contrasts.part)

# New formula for LMMs
form_joint <- tabulate.formula(~ experiment*part*condition + (part*condition|participant) + (experiment*part*condition|item))
form_joint %<>% mutate(block = replace(block, is.na(grouping), "fixed"))

# Check if models were computed already (delete the file to re-run)
if (file.exists("output/joint-stats.RDS")){
   # Load models from file
  models_joint <- readRDS("output/joint-stats.RDS")
} else {
  # Actually apply the modelling function
  models_joint <- map(comps$name, compute_models, formula = form_joint, data = data_joint, control = ctrl_params, specs = specs) %>%
    set_names(comps$name)
  # Backup epochs to the export folder
  saveRDS(models_joint, "output/joint-stats.RDS")}
```

\newpage

# References

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs custom-style="Bibliography"}
:::

\endgroup