### 2020-09-02 manuscript/manuscript.tex
render plots

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa7}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The title},
  pdfauthor={First Author1 \& Ernst-August Doelle1,2},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Title}
\keywords{keywords\newline\indent Word count: X}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {}%
  {\par}

\title{The title}
\author{First Author\textsuperscript{1} \& Ernst-August Doelle\textsuperscript{1,2}}
\date{}

\authornote{

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

The authors made the following contributions. First Author: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing.

Correspondence concerning this article should be addressed to First Author, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Wilhelm-Wundt-University\\\textsuperscript{2} Konstanz Business School}

\abstract{
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general context}.

Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
}

\begin{document}
\maketitle

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:exp1-output}Results of linear mixed-effects regression models for Experiment 1}

\footnotesize{

\begin{tabular}{lcccccc}
\toprule
 & \multicolumn{2}{c}{\textbf{P1}} & \multicolumn{2}{c}{\textbf{N1}} & \multicolumn{2}{c}{\textbf{N400}} \\
\cmidrule(r){2-3} \cmidrule(r){4-5} \cmidrule(r){6-7}
\textit{Type III} & \textit{F} (\textit{df}) & \textit{p} & \textit{F} (\textit{df}) & \textit{p} & \textit{F} (\textit{df}) & \textit{p}\\
\midrule
Part & 10.30 (2,   23.4) & .001 & 13.56 (2,   23.9) & <.001 & 31.61 (2,   24.4) & <.001\\
Insight & 0.87 (1, 5299.9) & .352 & 0.96 (1, 5341.1) & .327 & 21.83 (1, 5210.4) & <.001\\
Pt. × ins. & 2.37 (2, 4616.5) & .094 & 5.34 (2, 4949.3) & .005 & 11.16 (2, 5058.3) & <.001\\
\textit{Insight $-$  naive} & Est. [95\% CI] & \textit{p} & Est. [95\% CI] & \textit{p} & Est. [95\% CI] & \textit{p}\\ \midrule
Part 1 & 0.00 [-0.50, 0.49] & .991 & -0.17 [-0.62,  0.29] & .480 & -0.21 [-0.61, 0.18] & .292\\
Part 2 & -0.15 [-0.65, 0.35] & .553 & -0.66 [-1.12, -0.20] & .005 & 0.91 [ 0.51, 1.31] & <.001\\
Part 3 & 0.57 [ 0.08, 1.07] & .024 & 0.41 [-0.05,  0.87] & .080 & 0.99 [ 0.59, 1.39] & <.001\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Pt. = part, ins. = insight, est. = estimate, CI = confidence interval.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/exp1-output-1}

}

\caption{ }\label{fig:exp1-output}
\end{figure}

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:exp2-output}Results of linear mixed-effects regression models for Experiment 2}

\footnotesize{

\begin{tabular}{lcccccc}
\toprule
 & \multicolumn{2}{c}{\textbf{P1}} & \multicolumn{2}{c}{\textbf{N1}} & \multicolumn{2}{c}{\textbf{N400}} \\
\cmidrule(r){2-3} \cmidrule(r){4-5} \cmidrule(r){6-7}
\textit{Type III} & \textit{F} (\textit{df}) & \textit{p} & \textit{F} (\textit{df}) & \textit{p} & \textit{F} (\textit{df}) & \textit{p}\\
\midrule
Part & 4.71 (2,   22.6) & .020 & 15.69 (2,   23.9) & <.001 & 67.52 (2,   24.1) & <.001\\
Insight & 0.05 (1, 5221.8) & .826 & 1.24 (1, 5215.3) & .266 & 18.64 (1, 5203.6) & <.001\\
Pt. × ins. & 1.72 (2, 3556.7) & .180 & 1.33 (2, 4746.9) & .266 & 8.01 (2, 4652.4) & <.001\\
\textit{Insight $-$  naive} & Est. [95\% CI] & \textit{p} & Est. [95\% CI] & \textit{p} & Est. [95\% CI] & \textit{p}\\ \midrule
Part 1 & -0.17 [-0.72, 0.37] & .534 & -0.04 [-0.54, 0.45] & .858 & 0.02 [-0.44, 0.48] & .935\\
Part 2 & -0.32 [-0.87, 0.24] & .264 & -0.50 [-0.99, 0.00] & .050 & 1.31 [ 0.85, 1.77] & <.001\\
Part 3 & 0.38 [-0.18, 0.94] & .181 & 0.04 [-0.45, 0.54] & .864 & 0.45 [-0.01, 0.92] & .054\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Pt. = part, ins. = insight, est. = estimate, CI = confidence interval.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/exp2-output-1}

}

\caption{ }\label{fig:exp2-output}
\end{figure}

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:joint-output}Results of linear mixed-effects regression models for Experiments 1 and 2 combined}

\footnotesize{

\begin{tabular}{lcccccc}
\toprule
 & \multicolumn{2}{c}{\textbf{P1}} & \multicolumn{2}{c}{\textbf{N1}} & \multicolumn{2}{c}{\textbf{N400}} \\
\cmidrule(r){2-3} \cmidrule(r){4-5} \cmidrule(r){6-7}
\textit{Type III} & \textit{F} (\textit{df}) & \textit{p} & \textit{F} (\textit{df}) & \textit{p} & \textit{F} (\textit{df}) & \textit{p}\\
\midrule
Part & 11.83 (2,    47.0) & <.001 & 28.64 (2,    47.8) & <.001 & 91.32 (2,    48.4) & <.001\\
Insight & 0.35 (1, 10554.2) & .554 & 1.64 (1,   124.2) & .203 & 38.73 (1, 10511.3) & <.001\\
Experiment & 0.34 (1,    46.2) & .564 & 2.29 (1,    46.1) & .137 & 4.56 (1,    46.4) & .038\\
Pt. × ins. & 3.96 (2,  9094.4) & .019 & 5.59 (2,  9889.2) & .004 & 15.71 (2, 10020.2) & <.001\\
Pt. × exp. & 0.03 (2,    47.0) & .971 & 0.53 (2,    47.8) & .595 & 2.47 (2,    48.4) & .095\\
Ins. × exp. & 0.62 (1, 10530.2) & .432 & 0.04 (1, 10325.9) & .838 & 0.03 (1, 10567.6) & .867\\
Pt. × ins. × exp. & 0.02 (2,  9094.8) & .976 & 0.63 (2,  9889.5) & .533 & 2.64 (2, 10020.5) & .072\\
\textit{Insight $-$  naive} & Est. [95\% CI] & \textit{p} & Est. [95\% CI] & \textit{p} & Est. [95\% CI] & \textit{p}\\ \midrule
Part 1 & -0.07 [-0.44, 0.30] & .703 & -0.10 [-0.44,  0.24] & .563 & -0.10 [-0.41, 0.20] & .507\\
Part 2 & -0.22 [-0.59, 0.15] & .251 & -0.56 [-0.90, -0.21] & .002 & 1.09 [ 0.78, 1.39] & <.001\\
Part 3 & 0.49 [ 0.12, 0.86] & .010 & 0.25 [-0.10,  0.59] & .158 & 0.72 [ 0.41, 1.02] & <.001\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Pt. = part, ins. = insight, exp. = experiment, est. = estimate, CI = confidence interval.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{cslreferences}
\end{cslreferences}

\endgroup

\end{document}