### 2021-01-26 analysis/appendix_a.Rmd
add new appendix

---
title             : "Appendix"
shorttitle        : ""

author: 
  - name        : ""
    affiliation : ""

affiliation:
  - id          : ""
    institution : ""

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
numbersections    : true

documentclass     : "apa7"
classoption       : "man"
output            :
  papaja::apa6_pdf:
    latex_engine  : "xelatex"

header-includes:
  - \usepackage[export]{adjustbox}
  - \usepackage{makecell}
  - \renewcommand{\cellset}{\renewcommand{\arraystretch}{0.7}}
---

## Appendix A: Stimuli {.unnumbered}

\bigskip

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

```{r appendix, include=TRUE, warning=FALSE, message=FALSE, results="asis"}
# Load packages
library(papaja)
library(tidyverse)

# Print table of materials
read_tsv("materials/items.txt", col_types = cols()) %>%
  mutate(
    match_ger = paste0(match_noun_ger, ", ", match_verb_ger),
    match_en = paste0("{[", match_noun_en, ", ", match_verb_en, "]}"),
    nonmatch_ger = paste0(nonmatch_noun_ger, ", ", nonmatch_verb_ger),
    nonmatch_en = paste0("{[", nonmatch_noun_en, ", ", nonmatch_verb_en, "]}"),
  ) %>%
  transmute(
    img = paste0("\\includegraphics[valign=c, scale=0.23]{../materials/unfamiliar/", item, ".png}"),
    item = as.character(item),
    match = paste(match_ger, match_en, sep = "\n") %>%
      kableExtra::linebreak(align = "l", linebreaker = "\n"),
    nonmatch = paste(nonmatch_ger, nonmatch_en, sep = "\n") %>%
      kableExtra::linebreak(align = "l", linebreaker = "\n")
  ) %>%
  apa_table(
    booktabs = TRUE,
    col.names = c("Picture", "ID", "Matching keywords", "Non-matching keywords"),
    escape = FALSE,
    longtable = TRUE,
    caption = "Unfamiliar Object Stimuli\\smallskip",
    font_size = "footnotesize"
  )
```