### 2021-01-26 analysis/appendix_b.Rmd

---
title             : "Appendix"
shorttitle        : ""

author: 
  - name        : ""
    affiliation : ""

affiliation:
  - id          : ""
    institution : ""

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
numbersections    : true

documentclass     : "apa7"
classoption       : "man"
output            :
  papaja::apa6_pdf:
    latex_engine  : "xelatex"
---

## Appendix B: Linear Mixed-Effects Models {.unnumbered}

\noindent Linear mixed-effects models after model selection for Experiment 1:

\footnotesize

```{r exp1_formulas, include=TRUE}
# Load packages
library(purrr)

# Define function to print the formulas
print_formula <- function(model) cat(deparse(formula(model), width.cutoff = 75), sep = "\n")

# Formulas for models for Experiment 1
readRDS("analysis/export/exp1-models.RDS") %>% walk(~ print_formula (.x$model))
```

\normalsize

\noindent Linear mixed-effects models after model selection for Experiment 2:

\footnotesize

```{r exp2_formulas, include=TRUE}
# Formulas for models for Experiment 2
readRDS("analysis/export/exp2-models.RDS") %>% walk(~ print_formula (.x$model))
```

\normalsize

\noindent Linear mixed-effects models after model selection for Experiments 1 and 2 combined:

\footnotesize

```{r joint_formulas, include=TRUE}
# Formulas for joint models
readRDS("analysis/export/joint-models.RDS") %>% walk(~ print_formula (.x$model))
```

\normalsize