### 2022-02-04 analysis/manuscript.Rmd
Plot clusters

---
title             : "Instant effects of semantic information on visual perception"
shorttitle        : "Semantically informed perception"
author: 
  - name          : "Alexander Enge"
    affiliation   : "1,2"
    corresponding : yes
    address       : "Rudower Chaussee 18, 12489 Berlin, Germany"
    email         : "alexander.enge@hu-berlin.de"
  - name          : "Franziska Süß"
    affiliation   : "3"
  - name          : "Rasha Abdel Rahman"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Humboldt-Universität zu Berlin"
  - id            : "2"
    institution   : "Max Planck Institute for Human Cognitive and Brain Sciences"
  - id            : "3"
    institution   : "Fachhochschule des Mittelstands"
authornote: |
  \addORCIDlink{Alexander Enge}{0000-0003-0100-2297}

  \addORCIDlink{Rasha Abdel Rahman}{0000-0002-8438-1570}

  The authors declare no competing financial interests.
  The preprocessed data and analaysis code for this study are openly available at https://osf.io/uksbc/.
abstract: |
  Does our perception of an object change as soon as we discover what function it serves?
  This question matters not only when we encounter novel tools or gadgets in our everyday lives, but also pertains to the long-standing debate around the (im)penetrability of perception by higher cognitive capacities.
  Here, we showed human participants (*n* = 48) pictures of unfamiliar objects either together with words matching their function, leading to semantically informed perception, or together with non-matching words, resulting in naive perception.
  We measured event-related potentials (ERPs) to investigate at which stages in the visual processing hierarchy these two types of object perception differed from one another.
  We found that semantically informed as compared to naive perception was associated with larger amplitudes in the N170 component (150--200 ms) and reduced amplitudes in the N400 component (400--700 ms).
  When the same objects were presented once more without any information, the N400 effect persisted and we also observed enlarged amplitudes in the P1 component (100--150 ms) in response to objects for which semantically informed perception had taken place.
  Consistent with previous work, our results suggest that obtaining semantic information about previously unfamiliar objects alters aspects of their lower-level visual perception (P1 component), higher-level visual perception (N170 component), and semantic processing (N400 component).
  We demonstrate for the first time that these effects are instantaneous in the sense that they can be observed within the same trial as well as one trial after the information has been given, providing strong evidence that cognition affects perception.
keywords          : "objects, semantic knowledge, visual perception, event-related potentials"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
bibliography      : ["references.bib", "r-references.bib"]
csl               : "apa.csl"
documentclass     : "apa7"
classoption       : "man,donotrepeattitle"
fontsize          : "10pt"
output:
  papaja::apa6_pdf:
    latex_engine: xelatex
  papaja::apa6_docx: default
editor_options: 
  chunk_output_type: console
header-includes:
  - \geometry{a4paper}
  - \fancyheadoffset[R,L]{0pt}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage[bottom]{footmisc}
  - \interfootnotelinepenalty=10000
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\doublespacing}
  - \usepackage{makecell}
  - \renewcommand{\cellset}{\renewcommand{\arraystretch}{0.7}}
  - \captionsetup{font={stretch=1}, belowskip=15pt}
---

```{r, setup, include=FALSE}
# Load R packages
library(buildmer)
library(emmeans)
library(here)
library(knitr)
library(lme4)
library(lmerTest)
library(MASS)
library(papaja)
library(parallel)
library(reticulate)
library(tidyverse)
library(magrittr)

# Set global chunk options
opts_chunk$set(fig.width = 10, include = FALSE, out.width = "100%")

# Write R packages to bibliography
r_refs(here("analysis", "r-references.bib"), append = FALSE)

# What to (re-)run
run <- list(
  "preprocessing" = TRUE,
  "modeling" = TRUE
)

# Directory paths
data_dir <- here("data")
output_dir <- here("output")
report_dir <- here(output_dir, "html_reports")

# File paths
trials_file <- here(output_dir, "trials.csv")
models_file <- here(output_dir, "models.RDS")
clusters_file <- here(output_dir, "clusters.csv")
tfr_clusters_file <- here(output_dir, "tfr_clusters.csv")
```

```{r, preprocessing, eval=run$preprocessing}
# Process behavioral data
c(
  list.files(here(data_dir, "rt/exp1"), "*.txt", full.names = TRUE),
  list.files(here(data_dir, "rt/exp2"), "*.txt", full.names = TRUE)
) %>%
  # Process each log file
  map(function(log_file) {

    # Read the file
    read_tsv(
      log_file,
      col_types = cols(),
      locale = locale(encoding = "latin1")
    ) %>%
      # Recode some columns
      transmute(

        # Phase of the experiment
        phase = factor(Wdh,
          levels = c(211, 212, 213),
          labels = c("Pre-insight", "Insight", "Post-insight")
        ),

        # # Participant ID (unique for Experiments 1 and 2)
        # participant = ifelse(
        #   str_detect(log_file, "exp2"), str_c(VPNummer, "_exp2"), VPNummer
        # ),

        # Keyword conditions
        keywords = factor(Bed,
          levels = c("richtig", "falsch"),
          labels = c("Match", "Non-match")
        ),

        # Behavioral responses
        # 201: "I know what this is or have a strong assumption"  -> Recode as 4
        # 202: "I have an assumption what this is"                -> Recode as 3
        # 203: "I have rather no assumption what this is"         -> Recode as 2
        # 204: "I don't know what this is and have no assumption" -> Recode as 1
        response = 5 - (Tastencode - 200),

        # Reaction times
        rt = RT,
        log_rt = log(RT),

        # Item IDs
        item_id = factor(StimID)
      ) %>%
      # Remove filler items
      drop_na(keywords) -> log

    # Assign items to conditions based on manipulation (matching vs. non-
    # matching keywords) and the responses of the participants
    items_per_condition <- with(log, {
      list(

        # Informed condition: Matching keywords and positive response
        "Informed" = item_id[
          phase == "Insight" & keywords == "Match" & response %in% c(3, 4)
        ],

        # Naive condition: Non-matching keywords and negative response
        "Naive" = item_id[
          phase == "Insight" & keywords == "Non-match" & response %in% c(1, 2)
        ],

        # Exclude: Matching keywords and negative response
        "Exclude_informed" = item_id[
          phase == "Insight" & keywords == "Match" & response %in% c(1, 2)
        ],

        # Exclude: Non-matching keywords and positive response
        "Exclude_naive" = item_id[
          phase == "Insight" & keywords == "Non-match" & response %in% c(3, 4)
        ],

        # Exclude: Positive response *before* any info was presented
        "Exclude_known" = item_id[phase == "Pre-insight" & response == 4]
      )
    })

    # Assign these conditions to the trials of all three phases
    log$condition <- NA
    for (condition in names(items_per_condition)) {
      item_id <- items_per_condition[[condition]]
      log$condition[log$item_id %in% item_id] <- condition
    }

    # Sort columns
    log %>%
      select(item_id, phase, condition, response, rt, log_rt) %>%
      return()
  }) -> logs

# Import EEG pipeline from Python
pipeline <- import("pipeline")

# Get list of BrainVision EEG header files
vhdr_files <- c(
  list.files(here(data_dir, "eeg/exp1"), "*.vhdr", full.names = TRUE),
  list.files(here(data_dir, "eeg/exp2"), "*.vhdr", full.names = TRUE)
)

# Process EEG data
res <- pipeline$group_pipeline(
  vhdr_files = vhdr_files,
  log_files = logs,
  output_dir = output_dir,
  report_dir = report_dir,
  downsample_sfreq = 250.,
  bad_channels = NULL,
  ocular_correction = "fastica",
  highpass_freq = 0.1,
  lowpass_freq = 40.,
  triggers = c(221, 222),
  epochs_tmin = -0.5,
  epochs_tmax = 1.5,
  baseline = c(-0.2, 0.0),
  reject_peak_to_peak = 200.,
  reject_flat = 1.,
  components = list(
    "name" = c("P1", "N170", "N400"),
    "tmin" = c(0.1, 0.15, 0.4),
    "tmax" = c(0.15, 0.2, 0.7),
    "roi" = list(
      c("PO3", "PO4", "POz", "O1", "O2", "Oz"),
      c("P7", "P8", "PO7", "PO8", "PO9", "PO10"),
      c("C1", "C2", "Cz", "CP1", "CP2", "CPz")
    )
  ),
  average_by = c("phase/condition"),
  perform_tfr = TRUE,
  perm_contrasts = list(
    c("Pre-insight/Informed", "Pre-insight/Naive"),
    c("Insight/Informed", "Insight/Naive"),
    c("Post-insight/Informed", "Post-insight/Naive")
  ),
  n_jobs = 1
)
```

```{r, modeling, eval=run$modeling}
# Read single trial data
read_csv(trials_file, col_types = cols(
  participant_id = col_factor(),
  item_id = col_factor(),
  phase = col_factor(),
  condition = col_factor()
)) %>%
  # Mark unrealistically short RTs
  mutate(rt = ifelse(rt < 200, NA, rt)) -> trials

# Filter relevant conditions for the main analysis
trials %>%
  mutate(condition = factor(condition, levels = c("Informed", "Naive"))) %>%
  drop_na() -> trials_main

# Contrast coding for phase
cbind(
  c("Pre-insight" = -1, "Insight" = 1, "Post-insight" = 0),
  c("Pre-insight" = 0, "Insight" = -1, "Post-insight" = 1)
) %>%
  t() %>%
  ginv() -> contrasts(trials_main$phase)

# Contrast coding for condition
cbind(c("Informed" = 1, "Naive" = -1)) %>%
  t() %>%
  ginv() -> contrasts(trials_main$condition)

# Construct model formula
form <- ~ phase * condition +
  (phase * condition | participant_id) +
  (phase * condition | item_id)

# Fit linear mixed models
deps <- c("P1", "N170", "N400")
map(deps, function(dep, model_formula = form, model_data = trials_main) {

  # Modify the formula such that fixed effects are always retained
  model_formula %<>%
    tabulate.formula() %>%
    mutate(block = replace(block, is.na(grouping), "fixed"))

  # Start a CPU cluster so that multiple models can be fitted in parallel
  n_cores <- detectCores()
  cl <- makeCluster(n_cores)

  # Find the best fitting model (see Matuschek et al., 2017, *JML*)
  build <- buildmer(
    buildmerControl = buildmerControl(
      formula = model_formula,
      data = model_data,
      args = list("control" = lmerControl(
        optimizer = "bobyqa", optCtrl = list(maxfun = 1e6)
      )),
      direction = c("backward", "backward"),
      cl = cl,
      elim = LRTalpha(.20),
      calc.anova = TRUE,
      ddf = "Satterthwaite",
      dep = dep
    )
  )
  stopCluster(cl)

  # Compute marginal means and contrasts
  emm_options(lmer.df = "Satterthwaite", lmerTest.limit = Inf)
  specs <- pairwise ~ condition | phase
  em <- emmeans(build@model, specs, infer = TRUE)

  # Create list of returns
  list(
    "model" = build@model,
    "summary" = build@summary,
    "anova" = build@anova,
    "means" = em$emmeans,
    "contrasts" = em$contrasts
  ) %>%
    return()
}) %>%
  # Add ERP component names
  set_names(deps) -> models

# Save models
saveRDS(models, models_file)
```

```{r, table_1, include=TRUE}
# Read fitted models
models <- readRDS(models_file)

# View model summaries
walk(models, function(model) {
  print(formula(model$model))
  cat("\n")
  print(model$anova)
  cat("\n")
  print(model$contrasts)
  cat("\n\n")
})
```

```{r, clusters, include=TRUE, fig.height=8}
# Read results of cluster-based permutation tests for ERPs
read_csv(clusters_file, col_types = cols(
  contrast = col_factor(levels = c(
    "Pre-insight/Informed - Pre-insight/Naive",
    "Insight/Informed - Insight/Naive",
    "Post-insight/Informed - Post-insight/Naive"
  ))
)) -> clusters

# Plot thresholded cluster images
clusters %>%
  filter(p_val < .05) %>%
  ggplot(aes(x = time, y = channel, fill = cluster)) +
  facet_grid(~contrast, drop = FALSE) +
  geom_raster() +
  theme(legend.position = "top")
```

```{r, tfr, include=TRUE, fig.height=12}
# Read results of cluster-based permutation tests for TFR
read_csv(tfr_clusters_file, col_types = cols(
  contrast = col_factor(levels = c(
    "Pre-insight/Informed - Pre-insight/Naive",
    "Insight/Informed - Insight/Naive",
    "Post-insight/Informed - Post-insight/Naive"
  ))
)) -> tfr_clusters

# Plot thresholded cluster images for the insight phase
tfr_clusters %>%
  filter(contrast == "Insight/Informed - Insight/Naive" & p_val < .05) %>%
  ggplot(aes(x = time, y = freq, fill = cluster)) +
  facet_wrap(~channel) +
  geom_raster() +
  theme(legend.position = "top")

# Plot thresholded cluster images for the post-insight phase
tfr_clusters %>%
  filter(
    contrast == "Post-insight/Informed - Post-insight/Naive" & p_val < .05
  ) %>%
  ggplot(aes(x = time, y = freq, fill = cluster)) +
  facet_wrap(~channel) +
  geom_raster() +
  theme(legend.position = "top")
```

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup